Variation in gaze understanding across the life span: A process-level perspective

### 2023-08-02 paper/manuscript.Rmd
edit manuscript figures and corresponding saves

keywords          : "social-cognitive development, theory of mind, gaze cues, individual differences, cognitive modeling, lifespan"
wordcount         : "xxx max 8000"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes 
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
header-includes   :
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1}}
editor_options    : 
  markdown: 
    wrap: sentence
bibliography      : "../../references.bib"
---

```{r setup, include = FALSE}
library("tidyverse")
library("papaja")
library("kableExtra")
library("brms")

@@

**Funding:** This study was funded by the Max Planck Society for the Advancement of Science, a noncommercial, publicly financed scientific organization (no grant number).
We thank all the children, caregivers, and adults who participated in the study.

@@

While the emergence of gaze following has been well established, less is known about the developmental trajectory throughout childhood and adolescence.
One possibility is that our social-cognitive ability in question is fully developed once emerged in infancy.However, many cognitive abilities develop with age [e.g., working memory, @gathercole2004structure].
Similarly, visual processing appears to improve with age.
Therefore, children could potentially improve in gaze following, fine-tuning the performance of the already existing skill.

@@

We aimed at assessing the developmental trajectory of gaze understanding across the lifespan.
First, we were interested in how this ability changes with age: Is the ability to understand gaze fully developed once emerged in infancy and stays stable across the lifetime?
Or do we fine-tune our already existing ability in early adulthood?
Do we then potentially even notice an age decay later in adulthood?
Second, we were interested in individual differences across the age groups.
Does the variation between individuals decrease, the older they get?
Do all adults reach the same level of precision in gaze understanding?

@@

```{r lifespan_sample, include = T, out.width = "100%"}
lifespan_sample <- readRDS("../saves/lifespan_sample.rds")

lifespan_table <- lifespan_sample %>% 
  mutate(
    n = str_glue_data(., "{n_total} ({n_females} female)"), 
    `Age range` = str_glue_data(., "{age_min} - {age_max}"), 
  ) %>% 
  rename(
    `Age group` = age_group,
    `Age SD` = age_sd, 
    `Age mean` = age_mean
  ) %>% 
  select(`Age group`, n, `Age mean`, `Age range`, `Age SD`)

apa_table(lifespan_table)
```

@@

We collected data from a remote child, teenager and adult sample, reaching from 3 to 80 years of age.

@@

Each trial presented an agent standing in a window, watching a balloon (*i.e.*, target) falling to the ground.
The target fell behind a hedge, while the agent's gaze followed the target's trajectory: pupil and iris moved so that their center aligned with the target center.
In test trials, the target flight was covered so that participants could not see where the target landed.
Participant were tasked to locate the target by tracking the agent's gaze.
They could respond by touching on the screen.

@@

The outcome measure was imprecision, defined as the absolute difference between the target center and the x coordinate of the participant's click.
Target coordinates were randomly generated during runtime.
Each target bin, as well as all agents and target colors, occurred equally often and and did not appear in more than two consecutive trials.

(ref:figlab1) **Setup of the TANGO and Magnet tasks**. (A) TANGO: Gaze understanding task. The agent stands in a window with the target in front of them. A hedge grows and covers the target. The target falls to a random location on the ground. The agent's eyes track the movement of the target. (B) Magnet task: non-social vector estimation.

```{r fig1, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab1)"}
knitr::include_graphics("../figures/procedure.png")
```

@@

Children and teenagers received a personalized link to the study website.
Caregivers were asked to provide technical support whenever needed, while explicitly being reminded to not help their children in responding.
Webcam videos were recorded whenever consented and technically feasible, in order to monitor whether children and teenagers responded on their own.

Adults completed the online study unsupervised.

@@

To estimate the effect of age on gaze understanding, we fit GLMMs that make different assumptions about the developmental trajectory, modeling the relationship as linear, quadratic, cubic.
In addition, we applied a Gaussian Process model [@burkner2017brms] and examined nonlinearity.
The models predicted imprecision by age (continuous).
[^1] Imprecision was defined as the absolute click distance between the target center and the click X coordinate, scaled according to target widths, aggregated across trials, and modeled by a `lognormal` distribution.
We inspected the posterior distributions (mean and 95% Credible Interval (CrI)) for the age estimates and compared models using LOO and WAIC scores and model weights.

[^1]: Originally, our models implemented the following structure: `performance ~ age + symmetricPosition + trialNr + (1 + symmetricPosition + trialNr | subjID)`.
    However, the Gaussian Process model was computationally heavy.
    To circumvent issues of convergence and for practicability, we simplified the model structure to include only age as an effect.
    In the Supplements, you can find a comparison between the original and the here reported model structures.
    Essentially, the model predictions did not differ notably.

To further assess developmental change, we performed a Bayesian change-point analysis, using the package `RBeast` [@zhao2019detecting].
The function `beast.irreg` handled the irregular nature of our series, which did not have the same number of data points for each year in age.
We were interested in finding the most prominent, most likely change-points in our data, assuming a constant mean (i.e., a flat line) within each segment.
Additionally, we wanted to avoid "overreactions" to individual outlying data points.
Therefore, we restricted our search to a maximum number of five change-points and a minimum length of 9.1245 data points between change-points (corresponding to the product of mean age difference between consecutive points \* mean number of data points within 3 years \* 3).
We inspected the probability of number of change-points and the locations of these change-points (mean and 95% CrI).
See Supplements for change-point analyses with differently set parameters.

@@

(ref:figlab2) **Differentiation in gaze understanding**. Performance is measured as imprecision, i.e., the absolute distance between the target's center and the participant's click (averaged across trials). The unit of imprecision is counted in the width of the target, i.e., a participant with imprecision of 1 clicked on average one target width to the left or right of the true target center.

```{r fig2, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab2)"}
knitr::include_graphics(c("../figures/lifespan_plot.png"))
```

[TODO] report model comparison

We found a clear non-linear development in gaze understanding across the lifespan.
Early in childhood, children quickly enhanced their level of proficiency.
Performance peaked around early adulthood.
There was a minor decay in elderly adulthood.
Our Gaussian Process model indicated non-linearity: XXX.
The Bayesian change-point analysis revealed four major shifts in gaze understanding during the lifespan: We see a very rapid initial improvement in early childhood (3 change-points in rapid succession), followed by a long period of minor, very slow change with slightly increasing levels of imprecision toward the eldest in our sample.

@@

Three-year-olds were surprisingly inaccurate in their responses.
One possible explanation could be that they simply lacked the ability to complete the task, potentially due to issues in gaze-following.
Contrasting our results with previous findings on infant gaze-following, this explanation is unlikely.
A more likely explanation would be that children were able to follow the agent's gaze but struggled to translate this implicit understanding into active behavior.
Another point to keep in mind is that we used subtle eye movements as cues.
Many existing studies let the agents move eye and head in parallel, therefore establishing a confound with greater (head) movement.
Relying exclusively on the eye movement might be trickier for children than when presented with a combined eye and head orientation.
The performance of the youngest children seems more consistent with performance demands than with a failure in gaze following.

Regarding the sample of elderly adults, there is most likely a sampling bias present.
Elderly which know how to participate in *Prolific* studies certainly might show greater cognitive fitness compared to their offline counterparts.

Methodological challenges arise when trying to compare data across ages from qualitatively and quantitatively different study tasks.
We could circumvent these issues by applying the exact same task for the entire life span.

@@

Our lifespan study showed that gaze understanding develops throughout childhood, and variation between individuals appears in all age groups.
The TANGO has previously been proven reliable for capturing inter-individual differences in gaze understanding [@prein2023tango].
The variation between participants is thus likely genuine and not due to random noise.
Now, we aim to understand what explains the developmental change and the variation across participants on a process level.
We present a theory of gaze understanding that explains how children use all available information to make inferences about the agent's gaze and its attentional focus, which leads them to correctly identify the target position.
We formalize this inference process in a computational cognitive model.

@@

We seek to explain how participants solved the TANGO.
Our gaze model assumes that participants use available gaze information to infer the target location.
We quantify how well our model explains the gaze understanding process by comparing it to two alternative models that make different assumptions of which information children use and where they consequently click to locate the target.
Therefore, our modeling framework consists of three mutually exclusive models: (1) a gaze model, (2) a random guessing model, and (3) a center bias model.
We gauge which model can best explain our data by conducting a model comparison.

With our modeling framework, we can formalize whether a participant (A) uses the available gaze information at all and (B) how accurately they are in using this gaze information to locate the target.
Regarding developmental change, we hypothesize that a dual process could explain the development of gaze understanding: The older children are, the more likely they are to use gaze as a cue, and the more precise they become at using it to locate the target.

@@

@@

Our goal was to describe the inferential process by which participants locate the target.
We compared three computational cognitive models: (1) a gaze model, (2) a random guessing model, and (3) a center bias model.
All cognitive models and model comparisons were implemented in WebPPL [@goodman2014design].

The gaze model assumes participants use the agent's gaze as a cue to locate the target.[^2]
Participants are assumed to observe the current location of the pupil and estimate the center of the agent's eyeball.
Connecting these two point estimates as a line yields a gaze vector: the unique vector that extends from the center of the agent's eyeball through the center of the pupil to the attentional focal point.
This inferential process involves some uncertainty about the true center of the eyeball.
Participants are assumed to sample gaze vectors from a Normal distribution centered around the true gaze vector.
They are assumed to do this independently for the left and right eye and then integrate the information to estimate the target's location.

[^2]: This model mirrors the logic of the TANGO programming code.
    In the online experiment, we read out the center point coordinates of the target and the agent's eyeball (i.e., the SVG coordinates).
    We then calculate a line between these two points: this is our gaze vector.
    Now, knowing the eyeball radius, we calculate the point of intersection at which the gaze vector meets the eyeball boundary.
    Finally, the agent's pupil moves from the center of the eyeball along the gaze vector to the intersection point.
    This way, the agent is animated to "look at" the target.
    In the gaze model, we assume participants go through these steps in reverse order.

The geometry of sampling from a distribution around the gaze vector leads to an interesting, testable group-level prediction.
As the pupil location varies, a fixed amount of uncertainty around the gaze vector corresponds to a varying degree of uncertainty in the estimated target location.
When the agent directs their gaze toward the very left or right side, the distribution around the gaze vector from which participants sample is comparatively wider than when the agent gazes centrally to the ground in front of them.
For illustrative purposes, imagine a similar phenomenon: pointing a torch light to the ground.
When one directs the torchlight straight onto the ground, the light beam is clearly defined and smaller than when one points the torchlight further away from oneself.
Therefore, our gaze model predicts that our trials vary in difficulty: participants should be more imprecise in locating the target the further out it lands.
If our data matches the pattern of this model prediction, this can act as evidence for the gaze model.
Therefore, our gaze model provides a quantitative theory of gaze understanding, with testable model predictions.

To gauge the plausibility of our gaze model, we implemented two models that represent alternatives about how participants solve the TANGO.
Participants who are overall very imprecise in locating the target might be less likely to use the agent's gaze as a cue at all.
The alternative models, therefore, do not assume that participants make use of the gaze cue.
The first alternative model estimates how likely a participant engages in random guessing.
This behaves like a coin toss, sampling from a uniform distribution over all possible coordinates (i.e., 0 - 1920).
The second alternative model estimates how likely a participant shows a center bias: Participants' clicks could be drawn toward the screen center since the agent and the starting point of the balloon are located there.
This bias is implemented as sampling from a Normal distribution, with the center of the screen as the mean and one target width as the standard deviation.

Our three proposed models make different predictions about how participants' clicks will be distributed for different target locations.
We visualize and evaluate these differences using correlations between the model predictions and the data.
Furthermore, we evaluated these probabilistic models based on the marginal likelihood of the data under each model.
The pairwise ratio of marginal likelihoods for two models is also known as the Bayes Factor.
This factor quantifies the quality of a model's predictions by averaging over the possible values of the model's parameters weighted by the prior probabilities of those parameter values.
It can be used to estimate how much more likely the data under one model are compared to the other.
Bayes Factors implicitly consider model complexity (i.e., Bayesian Occam's razor): models with more parameters often have a broader prior distribution over parameters, which might weaken potential gains in predictive accuracy.
Details on models, including code to run the models, information about priors for parameter estimation, and Markov chain Monte Carlo settings, can be found in the associated online repository.

@@

@@

For the gaze model, we found a high similarity between the model predictions and the observed data across the entire age range.
The developmental trajectory that we observed in the data closely resembled the predictions of the model.
The model explained XXX% [TODO] of the variance in the data.

(ref:figlab3) **Gaze funnel**

```{r fig3, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab3)"}
knitr::include_graphics("../figures/gazefunnel_combined.png")
```

@@

Our findings from Study 1 showed individual differences and a developmental change in gaze understanding across the lifespan.
To answer what develops with age and how participants differ from one another, we presented a formal cognitive model of gaze understanding.
Participants are modeled to observe all available gaze information, integrate it from both the agent's eyes, and consequently arrive at the attentional focus of the agent: the target location.
We assume that children become increasingly sensitive to gaze cues but that the inference process remains the same.
By conducting model comparisons, we could rule out that participants' responses can be explained by random guessing or a center bias.

A big advantage of using a computational modeling framework is that it can disentangle where people's errors come from.
Our computational model can explain what varies between precise and imprecise individuals.
There are two sources of errors: (1) do you actually use the eye gaze information?
and (2) how accurate are you at estimating the focus point when you do pay attention to the agent's eyes?
This way, we can track developmental changes in gaze understanding in a more fine-tuned way.

@@

A limitation of our model is that we cannot disentangle how much of the participants' uncertainty comes from a noisy estimate of the agent's attentional focus and how much is due to imprecise clicking (e.g., wanting to click somewhere but experiencing motor issues at aiming, adding random noise to the click).
Initially, our proposed gaze model consisted of three components: (1) a perceptual component, whereby the participant produces a noisy observation of the agent's eye angle, (2) an inferential component, whereby the participant estimates the agent's point of attention based on the above perceptual component, and (3) a motor component, whereby the participant samples a click location from a distribution centered around the above estimate of the agent's attentional focus.
Simulation studies suggested it was difficult to disentangle the independent noise terms from these three components, so we simplified the gaze model to include only the inferential component.

@@

Our computational cognitive model assumes that the ability to engage in vector estimation is a crucial component of mastering gaze understanding.
In this study, we sought to experimentally isolate the physical vector estimation component.
In addition, we inquired whether there are any other cognitive processes outside the vector estimation that constitute gaze understanding.
We aimed to assess whether there are exclusively task-specific processes at hand or whether gaze understanding recruits a general social-cognitive ability that is shared among other social-cognitive tasks.

First, we aimed to isolate the vector estimation component of the gaze understanding task.
We designed a new non-social vector estimation task that shared all crucial design features of the gaze understanding task.
Second, we assessed children's social-cognitive abilities by administering a ToM task battery, comprising four tasks from the ToM scale by Wellman and Liu (@wellman2004scaling) and two additional perspective-taking tasks [@flavell1981development; @flavell1981younga].

Our reasoning was that the gaze understanding task shares task demands with the non-social vector estimation task, while it shares its social context with the ToM tasks.
This way, we can disentangle what components comprise gaze understanding.

Task design, data collection, and sample sizes were pre-registered: <https://osf.io/xsqkt>.
The study design and procedure obtained ethical clearance by the MPG Ethics commission Munich, Germany, falling under a packaged ethics application (Appl. No. 2021_45), and was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology.
The research adheres to the legal requirements of psychological research with children in Germany.
Data were collected between February and March 2023.

>>>>

keywords          : "social-cognitive development, theory of mind, gaze cues, individual differences, cognitive modeling, lifespan"
wordcount         : "xxx max 8000"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes 
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
header-includes   :
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1}}
editor_options    : 
  markdown: 
    wrap: sentence
bibliography      : "../../references.bib"
bibliography: references.bib
---

```{r setup, include = FALSE}
library("tidyverse")
library("papaja")
library("kableExtra")
library("brms")
library("janitor")

@@

**Funding:** This study was funded by the Max Planck Society for the Advancement of Science, a noncommercial, publicly financed scientific organization (no grant number). Manuel Bohn was supported by a Jacobs Foundation Research Fellowship. We thank all the children, caregivers, and adults who participated in the study.

@@

While the emergence of gaze following has been well established, less is known about the developmental trajectory throughout childhood and adolescence. One possibility is that our social-cognitive ability in question is fully developed once emerged in infancy. However, many cognitive abilities develop with age [e.g., working memory, @gathercole2004structure]. Similarly, visual processing appears to improve with age. Therefore, children could potentially improve in gaze following, fine-tuning the performance of the already existing skill.

@@

We aimed to assess the developmental trajectory of gaze understanding across the lifespan. First, we were interested in how this ability changes with age: Is the ability to understand gaze fully developed once emerged in infancy and stays stable across the lifetime? Or do we fine-tune our already existing ability in early adulthood? Do we then potentially even notice an age decay later in adulthood? Second, we were interested in individual differences across the age groups. Does the variation between individuals decrease, the older they get? Do all adults reach the same level of precision in gaze understanding?

@@

```{r lifespan_sample, include = T, out.width = "100%"}
lifespan_sample <- readRDS("../saves/lifespan_sample.rds")
```

@@

We collected data from a remote child, teenager and adult sample, reaching from 3 to 80 years of age (see Supplements for further details).

@@

Each trial presented an agent standing in a window, watching a balloon (*i.e.*, target) falling to the ground. The target fell behind a hedge while the agent's gaze followed the target's trajectory: pupil and iris moved so that their center aligned with the target center. In test trials, the target flight was covered so that participants could not see where the target landed. Participants were tasked to locate the target by tracking the agent's gaze. They could respond by touching on the screen.

@@

The outcome measure was imprecision, defined as the absolute difference between the target (i.e., balloon) center and the x coordinate of the participant's click. The full screen width was divided into ten bins. Within each bin, exact target coordinates were randomly generated during runtime. Each target bin, as well as all agents and target colors, occurred equally often and did not appear in more than two consecutive trials.

@@

Children and teenagers received a personalized link to the study website. Caregivers were asked to provide technical support whenever needed, while explicitly being reminded not to help their children in responding. Webcam videos were recorded whenever consented and technically feasible in order to monitor whether children and teenagers responded on their own. Adults completed the online study unsupervised.

@@

To estimate the effect of age on gaze understanding, we fit GLMMs that make different assumptions about the developmental trajectory, modeling the relationship as linear, quadratic, or cubic. In addition, we applied a Gaussian Process model [@burkner2017brms] and examined nonlinearity. The models predicted imprecision by age (continuous), aggregated across trials and modeled as a `lognormal` distribution.[^1] The unit of imprecision was counted in target width, i.e., a participant with imprecision of 1 clicked on average one target width to the left or right of the true target center. We inspected the posterior distributions (mean and 95% Credible Interval (CrI)) for the age estimates and compared models using leave-one-out cross-validation (LOO), the widely applicable information criterion (WAIC), and model weights [@vehtari2017practicala].

[^1]: Originally, our models were fitted on a trial-by-trial basis and implemented the following structure: `performance ~ age + symmetricPosition + trialNr + (1 + symmetricPosition + trialNr | subjID)`. However, the Gaussian Process model was computationally heavy. To circumvent issues of convergence and for practicability, we simplified the model structure, aggregated data on a subject level, and included only age as an effect. In the Supplements, you can find a comparison between the original and the here reported model structures. Essentially, the model predictions did not differ notably.

Confirming that the developmental change was non-linear, we performed a Bayesian change point analysis, using the package `RBeast` [@zhao2019detecting]. The function `beast.irreg` handled the irregular nature of our series, which did not have the same number of data points for each year in age. We were interested in finding the most prominent, most likely change points in our data, assuming a constant mean (i.e., a flat line, zero degree polynomial) within each segment. To avoid "overreactions" to individual outlying data points, we constrained the model to have minimally 10 data points between two change points (i.e., corresponding to half of the data points we collected per adult decade). We inspected the probability of number of change points and the locations of these change points (mean and 95% CrI).[^2]

[^2]: In a supplementary analysis, we varied the parameters of our changepoint analysis. We modified the number of allowed change points, the minimum number of data points between change points, and the polynomial order. When we allowed more explorative room (i.e., greater nr of change points, smaller minimum nr of data points between change points, higher polynomial order), the models became more sensitive and added more fine-grained change points. The exact location of the change points varied slightly. However, the overall interpretation stayed the same, fitting our initial visual inspection. While early childhood was characterized by much change, adults showed a relatively stable level of imprecision. There was a minor change in that elderly adults became slightly more imprecise again. See Supplements for further detail.

@@

(ref:figlab1) **Differentiation in gaze understanding**. Developmental trajectory in gaze understanding across the lifespan. Performance is measured as imprecision, i.e., the absolute distance between the target's center and the participant's click (averaged across trials). The unit of imprecision is counted in the width of the target, i.e., a participant with imprecision of 1 clicked on average one target width to the left or right of the true target center. Grey dots show the mean performance for each subject averaged across trials. Blue lines show 100 draws from the expectation of the posterior predictive distribution (i.e, conditional expectation) of the Gaussian Process model, with its mean predicted developmental trajectory as a solid black line. Vertical, black, dashed lines show the location of the most prominent change points.

```{r fig1, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab2)"}
knitr::include_graphics(c("../figures/lifespan_all.png"))
```

```{r lifespan_results, include=FALSE}
lifespan_variation <- readRDS("../saves/lifespan_variation_all.rds")
lifespan_mcomparison <- readRDS("../saves/lifespan_mcomparison.rds") %>% mutate_if(is.numeric, round, 2)
lifespan_mgaussianprocess <- readRDS("../saves/lifespan_mgaussianprocess.rds")
lifespan_beast <- readRDS("../saves/lifespan_ibeast.rds")
lifespan_cps <- readRDS("../saves/lifespan_ibeast_cps.rds") %>% round(2)

lifespan_gp <- summary(lifespan_mgaussianprocess)$gp %>% 
  rownames_to_column("param") %>% 
  clean_names()
```

High levels of variation pointed to individual differences in all age groups (overall imprecision mean = `r lifespan_variation$imprecision_mean`, sd = `r lifespan_variation$imprecision_sd`, range = [`r lifespan_variation$imprecision_min` - `r lifespan_variation$imprecision_max`]). For example, there were some children that were more accurate than the average adult.

In our model comparison, we found clear evidence for a non-linear development in gaze understanding across the lifespan. Compared against a linear, quadratic, and cubic model, the Gaussian Process model showed the best model fit, demonstrated by the greatest model weight and smallest WAIC values. For the imprecision in gaze understanding, the standard deviation (*b* = `r lifespan_gp$estimate[lifespan_gp$param == "sdgp(gpage_centered)"]`, 95% CrI [`r lifespan_gp$l_95_percent_ci[lifespan_gp$param == "sdgp(gpage_centered)"]`; `r lifespan_gp$u_95_percent_ci[lifespan_gp$param == "sdgp(gpage_centered)"]`]) and visual inspection indicated nonlinearity (see Figure \@ref(fig:fig2)).

Going one step further, we investigated the most prominent change points in the data. The Bayesian change point analysis revealed `r lifespan_beast$trend$ncp %>% round(1)` (with `r lifespan_beast$trend$ncpPr[lifespan_beast$trend$ncp %>% round(1) + 1] * 100 %>% round(2)`% probability) major shifts in gaze understanding during the lifespan. The change points occurred at the following ages: `r lifespan_cps$cp[1]` years (95% CrI [`r lifespan_cps$lowerCrI[1]`; `r lifespan_cps$upperCrI[1]`]); followed by `r lifespan_cps$cp[2]` years (95% CrI [`r lifespan_cps$lowerCrI[2]`; `r lifespan_cps$upperCrI[2]`]); followed by `r lifespan_cps$cp[3]` years (95% CrI [`r lifespan_cps$lowerCrI[3]`; `r lifespan_cps$upperCrI[3]`]); and finally at `r lifespan_cps$cp[4]` years (95% CrI [`r lifespan_cps$lowerCrI[4]`; `r lifespan_cps$upperCrI[4]`]). In other words: we found a very rapid initial improvement in early childhood (three change points in rapid succession), followed by a long period of minor, very slow change with slightly increasing levels of imprecision toward the eldest in our sample.

@@

We investigated the shape of change in gaze understanding across the lifespan. By applying the exact same task for the entire life span, we could directly compare gaze understanding in all ages. Therefore, we circumvented methodological challenges that often arise when trying to compare data across ages from qualitatively and quantitatively different study tasks.

We found a non-linear developmental trajectory in gaze understanding: Early in childhood, children quickly enhanced their level of proficiency. Performance peaked (i.e., imprecision was lowest) around early adulthood, while there was a minor decay in later adulthood. This is consistent with the view that we fine-tune our existing gaze understanding ability after the first emergence in early childhood. Furthermore, we observed individual differences in all age groups. While variation was highest in the three- and four-year-olds, it remained relatively stable across the lifespan.

Previous studies found that children start to follow gaze in the second half of their first year of life [@moll200412]. In our sample, three-year-olds were still rather imprecise in their gaze understanding ability. How can we explain this divergence? First of all, we used subtle eye movements as cues. Many existing studies let the agents move eye and head in parallel [@behne2005oneyearolds; @povinelli1997exploitation], therefore establishing a confound with greater (head) movement. Relying exclusively on eye movements might be more difficult for children than presenting them with a combined eye and head orientation [@carpenter1998social]. Furthermore, our study required participants to (1) precisely follow an agent's gaze, (2) interpret this as a cue, and then (3) make use of this cue to guide their own behavior (i.e., touching the screen at the cued location). We measured an active location choice instead of an attention orientation. It is conceivable that three-year-olds followed the agent's gaze but were still learning to translate this understanding into precise, active behavior.

Regarding the sample of elderly adults, we expect a sampling bias to be present [@bethlehem2010selection; @gosling2004should; @remillard2014systematic]. First, certainly not all older people have working a high-speed Internet connection or are knowledgeable and trained in its use. Second, it takes a certain amount of independence and motivation to participate in *Prolific* studies. The elderly who know how to participate in *Prolific* studies certainly might show greater cognitive fitness and flexibility compared to their offline counterparts. Therefore, a representative sample might show a greater age decline in gaze understanding compared to our reported sample. In addition, it is noteworthy that older people might be more likely to suffer from visual impairments. Even though we filtered participants to only include normal- to correct-to-normal vision, we cannot guarantee that our participants showed no symptoms of reduced vision.

@@

Our lifespan study showed that gaze understanding develops throughout childhood, and variation between individuals appears in all age groups. The TANGO has previously been shown to be reliably capture inter-individual differences in gaze understanding [@prein2023tango]. The variation between participants is thus likely genuine and not due to random noise. Now, we aim to understand what explains the developmental change and the variation across participants on a process level. We present a theory of gaze understanding that explains how children process the available information (i.e., the agent's eyes) to make inferences about the agent's gaze and its attentional focus, which leads them to identify the target position. We formalize this inference process in a computational cognitive model.

@@

We seek to explain how participants solved the TANGO. Our gaze model assumes that participants use available gaze information to infer the target location.

@@

## Computational model

The gaze model assumes participants use the agent's gaze as a cue to locate the target.[^3] Participants are assumed to observe the current location of the pupil and estimate the center of the agent's eyeball. Connecting these two point estimates as a line yields a gaze vector: the unique vector that extends from the center of the agent's eyeball through the center of the pupil to the attentional focal point. Participants are assumed to sample gaze vectors from a Normal distribution centered around the true gaze vector. They are assumed to do this independently for the left and right eye and then integrate the information to estimate the target's location. This inferential process involves some uncertainty about the true gaze vector.

[^3]: This model mirrors the logic of the TANGO programming code. In the online experiment, we read out the center point coordinates of the target and the agent's eyeball (i.e., the SVG coordinates). We then calculate a line between these two points: this is our gaze vector. Now, knowing the eyeball radius, we calculate the point of intersection at which the gaze vector meets the eyeball boundary. Finally, the agent's pupil moves from the center of the eyeball along the gaze vector to the intersection point. This way, the agent is animated to "look at" the target. In the gaze model, we assume participants go through these steps in reverse order.

The geometry of sampling from a distribution around the gaze vector leads to an interesting, testable group-level prediction. As the pupil location varies, a fixed amount of uncertainty around the gaze vector corresponds to a varying degree of uncertainty in the estimated target location. When the agent directs their gaze toward the very left or right side, the distribution around the gaze vector from which participants sample is comparatively wider than when the agent gazes centrally to the ground in front of them. For illustrative purposes, imagine a similar phenomenon: pointing a torch light to the ground. When one directs the torchlight straight onto the ground, the light beam is clearly defined and smaller than when one points the torchlight further away from oneself. Therefore, our gaze model predicts that our trials vary in difficulty: participants should be more imprecise in locating the target the further out it lands.[^4] If our data matches the pattern of this model prediction, this can act as evidence for the gaze model. Therefore, our gaze model provides a quantitative theory of gaze understanding, with testable model predictions.

[^4]: In our screen-based study, this effect should decrease again towards the most outward sides. Since the computer screen has a natural border, trials in which the target lands furthest out to the left/right become slightly easier again. In these cases, the uncertainty about the true gaze vector is - practically - only on the inner side (facing the center) of the screen. The natural border of the screen limits where participants can respond. In another adult sample with more trials, we could recover this pattern. For further elaboration, see Supplements.

@@

Our goal was to describe the inferential process of gaze understanding. We quantified how well our model explained the gaze understanding process by comparing it to two alternative models that make different assumptions of which information children use and where they consequently click to locate the target. Our modeling framework consisted of three mutually exclusive models: (1) a gaze model, (2) a random guessing model, and (3) a center bias model. We gauged which model can best explain our data by conducting model comparisons. All cognitive models and model comparisons were implemented in WebPPL [@goodman2014design].

To gauge the plausibility of our gaze model, we implemented two models that represent alternatives about how participants solve the TANGO. Participants who were overall very imprecise in locating the target might be less likely to use the agent's gaze as a cue at all. The alternative models, therefore, do not assume that participants made use of the gaze cue. The first alternative model assumed participants were randomly guessing. This was implemented as sampling from a uniform distribution over all possible coordinates (i.e., 0 - 1920). The second alternative model assumed the participants were biased towards the screen center: Participants could be drawn toward the screen center since the agent and the starting point of the balloon were located there. This bias was implemented as sampling from a Normal distribution, with the center of the screen as the mean and one target width as the standard deviation.

Our three proposed models made different predictions about how participants' clicks would be distributed for different target locations. We visualized and evaluated these differences using correlations between the model predictions and the data. Furthermore, we evaluated these probabilistic models based on the marginal likelihood of the data under each model. The pairwise ratio of marginal likelihoods for two models is also known as the Bayes Factor. This factor quantifies the quality of a model's predictions by averaging over the possible values of the model's parameters weighted by the prior probabilities of those parameter values. It can be used to estimate how much more likely the data under one model are compared to the other. Bayes Factors implicitly consider model complexity (i.e., Bayesian Occam's razor): models with more parameters often have a broader prior distribution over parameters, which might weaken potential gains in predictive accuracy. Details on models, including code to run the models, information about priors for parameter estimation, and Markov chain Monte Carlo settings, can be found in the associated online repository.

[TODO: add more infos on priors & parameters]

@@

We found clear support for our gaze model, both in children as well as adults.

@@

For the gaze model, we found a high similarity between the model predictions and the observed data across the entire age range. The developmental trajectory that we observed in the data closely resembled the predictions of the model: the correlation between the data and the model predictions was XXX [TODO]. The model explained XXX% [TODO] of the variance in the data.

[TODO: report prior]

(ref:figlab2) **Gaze model**

```{r fig2, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab3)"}
knitr::include_graphics("../figures/gazemodel_plot.png")
```

@@

Our findings from Study 1 showed individual differences and a developmental change in gaze understanding across the lifespan. To answer what develops with age and how participants differ from one another, we presented a formal cognitive model of gaze understanding. Participants are modeled to observe all available gaze information, integrate it from both the agent's eyes, and consequently arrive at the attentional focus of the agent: the target location. We assume the basic process of gaze understanding to be the same across the lifespan, though individuals become increasingly precise with age. By conducting model comparisons, we could rule out that participants' responses can be explained by random guessing or a center bias.

@@

A limitation of our model is that we cannot disentangle how much of the participants' uncertainty comes from a noisy estimate of the agent's attentional focus and how much is due to imprecise clicking (e.g., wanting to click somewhere but experiencing motor issues at aiming, adding random noise to the click).

A critical feature of our model is that it assumes gaze understanding to rely on vector estimations. In other words, subjects are modeled to calculate gaze vectors based on the location of an agent's pupil. This way, they can infer the agent's attentional focus. Even though this vector estimation component is a rather physical, geometrical calculation, it still happens in a social context. In the first place, one must interpret the agent's eyes as a relevant social stimulus. Therefore, our computational model describes gaze understanding as a particular form of vector estimation in a social context.

@@

## Perspective-taking:

[@erle2017grounded]

-   differentiation between visuospatial perspective-taking and psychological (i.e., cognitive and affective) perspective-taking

-   unclear boundaries between ToM & perspective-taking, many correlational studies, not so many causation "terminological confusion"

-   persepctive-taking: stepping into somebody else's shoes. recognizing the other's thoughts, feelings, mental states (=\> ToM!!)

    -

        1)  attesting mental states to another social agent

    -

        2)  other's states can be ≠ to own

    -

        3)  put aside egocentric bias

-   little is known about underlying psychological mechanisms of perspective-taking

-   visuospatial perspective taking: mentally transforming yourself and your body into the physical location of another agent. judging differences between one's own egocentric point of view an that of another person

    -   level 1: visibility of objects from certain view point. can operate independently from another's frame of reference

    -   level 2: imagine how world looks like for another person "climbing into the skin of another person". literally rotating body schema (embodied) "transposing body schema into the target's position"

-   clinical populations show deficit in empathy & visuospatial perspective-taking

-   idea that psychological & visuospatial perspective-taking share common simulation-based mechanism

-   studies show that: "Taking another's perspective led participants to adopt the thoughts of the target person more strongly (Experiments 1--3) and increased the perceived similarity of that person to the self (Experiment 4) and participants' liking of that person (Experiment 5). These effects were independent of task difficulty (Experiment 2), and only present during trials where an embodied transformation happened (i.e., at high angular disparities; Experiment 3)." "not only affects social-cognitive outcomes, but also empathic perspective-taking" =\> "shared mechanism of all kinds of perspective-taking"

-   non-social perspective-taking task: display empty chair. spatial perspective-taking. lead to "diametrically different results"

[@birch2017perspectives]

-   "Specifically, we contend that these processes (a) are partially innate, (b) develop over time and can be honed through experience, and (c) lie on a continuum with some individuals being better than others."

-   Apperly & Buuterfill 2009: System 1 process social information relatively quickly & effortlessly, clear limitations, inflexible. System 2 explicit reasoning, involves language, inhibitory control, working memory, cognitive flexibility, reasoning, planning etc.

-   chapter assumes partial innateness, develop over time, can be honed through experience, lie on a continuum with some individuals being "Better" than others

-   children use gaze as indicators of others' mental states from an early age

-   individual differences:

    -   children differ in onset of milestones, degree to which they can and do use ToM in daily life. vary in frequency / propensity (connected to motivation?) and accuracy

    -   ind diff to "illustrate important function ToM serves in navigating the social world". for which processes involved, contribution of other skills & environmental scaffolding

-   curse of knowledge: U shaped curve. hindsight bias & tom correlated, but not inhibitory control as mediator
